ManyEcoEvo: an R package for working with the "Same Data, Different Analysts" Project in Ecology and Evolutionary Biology, and other many-analyst style data

### 2025-11-19 manuscript/manuscript/manuscript.Rmd
Update year from 2023 --> 2025, and switch to published many analysts paper citation.

---
# Example from https://joss.readthedocs.io/en/latest/submitting.html
title: 'ManyEcoEvo: an R package for working with the "Same Data, Different Analysts" Project in Ecology and Evolutionary Biology, and other many-analyst style data'
tags:
  - R
  - meta-analysis
  - metaresearch
  - ecology
  - evolutionary biology
  - many-analysts
authors:
  - name: Elliot Gould
    orcid: 0000-0003-0872-7098
    affiliation: "1" 
  - name: Hannah S. Fraser
    orcid: 0000-0000-0000-0000
    affiliation: "1"
  - name: Shinichi Nakagawa
    orcid: 0000-0000-0000-0000
    affiliation: "2"
  - name: Timothy H. Parker
    orcid: 0000-0000-0000-0000
    affiliation: "3"
affiliations:
 - name: The University of Melbourne
   index: 1
 - name: The University of New South Wales
   index: 2
 - name: Whitman College
   index: 3
citation_author: Gould and Fraser et al. 
date: "`r Sys.Date()`"
year: 2023
bibliography: paper.bib
output: 
  # word_document: default
  rticles::joss_article:
    journal: JOSS
csl: apa.csl
---

@@

*ManyEcoEvo* is an R package [@RCoreTeam2022] designed to *facilitate the analysis of* an existing many-analyst study in ecology and evolutionary biology, the *ManyEcoEvo Project* [@Gould2025]. The purpose of *ManyEcoEvo* is to reproducibly implement the analysis and presentation of results for Gould *et al.* [-@gould2023] such that the 246 analysts as well as other interested researchers may independently reproduce the study's findings. The package also aims to allow re-analyses and replications of Gould *et al.* [-@gould2023], as well as facilitating the analysis of similar many-analyst style datasets.

The ManyEcoEvo project is a many-analyst study that aimed to empirically explore the variation in effect sizes and model predictions generated by analytical decisions of different researchers within ecology and evolutionary biology [@gould2023]. 174 analyst teams, comprising 246 analysts, independently investigated the answers to two pre-specified research questions using two previously unpublished datasets, one from conservation ecology, and one from evolutionary biology. Analysts investigating the evolutionary biology dataset examined the effect of sibling number and nestling growth of blue tits (*Cyanistes caeruleus*), while analysts investigating the conservation ecology dataset examined the relationship between grass cover and *Eucalyptus* tree seedling recruitment. The ManyEcoEvo project quantified the variation in results among analyses using meta-analysis and additionally sought to understand how features of studies such as, variable selection, random effects structures, qualitative and numeric ratings of the methods by peer-reviewers, influenced deviation from the meta-analytic mean.

All raw data sets and metadata provided to independent teams for analysis are publicly archived at <https://osf.io/qjzby> and <https://osf.io/hdv8m>, respectively. Analyst survey responses and results are provided in the package GitHub repository (<https://github.com/egouldo/ManyEcoEvo/>), which is also archived on Zenodo (<https://doi.org/10.5281/zenodo.10046153>). *ManyEcoEvo* contains tools for data-cleaning and processing tasks encountered when working with many-analyst style data, functions for implementing and reproducing the analysis described inGould *et al.* [-@gould2023], establishes a structured workflow that facilitates re-analyses and replications of Gould *et al.* [-@gould2023] as well as implementation for new many-analyst studies, including a planned extension of the ManyEcoEvo project.

@@

We provide a suite of miscellaneous functions for data cleaning and processing tasks that we expect to be common to the sorts of data collected by many-analyst style studies. This was the most time and effort-intensive aspect of the analysis in Gould *et al.* [-@gould2023], and we hope these functions will ease the burden on future researchers conducting many-analyst studies.

*ManyEcoEvo* provides custom functions to calculate standardised effect sizes in the form of Fishers' Z correlation scores in preparation for meta-analysis. While some existing functions exist within other packages [e.g. `DescTools::FishersZ()` @DescTools, `DiagTest3Grp::FisherZ.var()` @luo2012diagtest3grp], the method for calculating variance of the standardised effect size in Gould *et al.* [-@gould2023] differed slightly, and we wished to avoid reliance on too many external packages [@wickham2023r, Chapter 10.1]. Should *ManyEcoEvo* users wish to deviate from the effect-size standardisation method used in Gould *et al.* [-@gould2023], they are free to use their own functions provided by other software or custom functions [@harrer2021, see chapter 3 for a range of R code to calculate various effect sizes].

@@

<!-- We acknowledge the contributions of all 246 analysts and co-authors of Gould *et al. [-@gould2023]* please see the original manuscript for a full list of contributors to the ManyEcoEvo project. -->

>>>>

---
# Example from https://joss.readthedocs.io/en/latest/submitting.html
title: 'ManyEcoEvo: an R package for working with the "Same Data, Different Analysts" Project in Ecology and Evolutionary Biology, and other many-analyst style data'
tags:
  - R
  - meta-analysis
  - metaresearch
  - ecology
  - evolutionary biology
  - many-analysts
authors:
  - name: Elliot Gould
    orcid: 0000-0003-0872-7098
    affiliation: "1" 
  - name: Hannah S. Fraser
    orcid: 0000-0000-0000-0000
    affiliation: "1"
  - name: Shinichi Nakagawa
    orcid: 0000-0000-0000-0000
    affiliation: "2"
  - name: Timothy H. Parker
    orcid: 0000-0000-0000-0000
    affiliation: "3"
affiliations:
 - name: The University of Melbourne
   index: 1
 - name: The University of New South Wales
   index: 2
 - name: Whitman College
   index: 3
citation_author: Gould and Fraser et al. 
date: "`r Sys.Date()`"
year: 2025
bibliography: paper.bib
output: 
  # word_document: default
  rticles::joss_article:
    journal: JOSS
csl: apa.csl
---

@@

*ManyEcoEvo* is an R package [@RCoreTeam2022] designed to *facilitate the analysis of* an existing many-analyst study in ecology and evolutionary biology, the *ManyEcoEvo Project* [@Gould2025]. The purpose of *ManyEcoEvo* is to reproducibly implement the analysis and presentation of results for Gould *et al.* [-Gould2025] such that the 246 analysts as well as other interested researchers may independently reproduce the study's findings. The package also aims to allow re-analyses and replications of Gould *et al.* [-Gould2025], as well as facilitating the analysis of similar many-analyst style datasets.

The ManyEcoEvo project is a many-analyst study that aimed to empirically explore the variation in effect sizes and model predictions generated by analytical decisions of different researchers within ecology and evolutionary biology [Gould2025]. 174 analyst teams, comprising 246 analysts, independently investigated the answers to two pre-specified research questions using two previously unpublished datasets, one from conservation ecology, and one from evolutionary biology. Analysts investigating the evolutionary biology dataset examined the effect of sibling number and nestling growth of blue tits (*Cyanistes caeruleus*), while analysts investigating the conservation ecology dataset examined the relationship between grass cover and *Eucalyptus* tree seedling recruitment. The ManyEcoEvo project quantified the variation in results among analyses using meta-analysis and additionally sought to understand how features of studies such as, variable selection, random effects structures, qualitative and numeric ratings of the methods by peer-reviewers, influenced deviation from the meta-analytic mean.

All raw data sets and metadata provided to independent teams for analysis are publicly archived at <https://osf.io/qjzby> and <https://osf.io/hdv8m>, respectively. Analyst survey responses and results are provided in the package GitHub repository (<https://github.com/egouldo/ManyEcoEvo/>), which is also archived on Zenodo (<https://doi.org/10.5281/zenodo.10046153>). *ManyEcoEvo* contains tools for data-cleaning and processing tasks encountered when working with many-analyst style data, functions for implementing and reproducing the analysis described inGould *et al.* [-Gould2025], establishes a structured workflow that facilitates re-analyses and replications of Gould *et al.* [-Gould2025] as well as implementation for new many-analyst studies, including a planned extension of the ManyEcoEvo project.

@@

We provide a suite of miscellaneous functions for data cleaning and processing tasks that we expect to be common to the sorts of data collected by many-analyst style studies. This was the most time and effort-intensive aspect of the analysis in Gould *et al.* [-Gould2025], and we hope these functions will ease the burden on future researchers conducting many-analyst studies.

*ManyEcoEvo* provides custom functions to calculate standardised effect sizes in the form of Fishers' Z correlation scores in preparation for meta-analysis. While some existing functions exist within other packages [e.g. `DescTools::FishersZ()` @DescTools, `DiagTest3Grp::FisherZ.var()` @luo2012diagtest3grp], the method for calculating variance of the standardised effect size in Gould *et al.* [-Gould2025] differed slightly, and we wished to avoid reliance on too many external packages [@wickham2023r, Chapter 10.1]. Should *ManyEcoEvo* users wish to deviate from the effect-size standardisation method used in Gould *et al.* [-Gould2025], they are free to use their own functions provided by other software or custom functions [@harrer2021, see chapter 3 for a range of R code to calculate various effect sizes].

@@

<!-- We acknowledge the contributions of all 246 analysts and co-authors of Gould *et al. [-Gould2025]* please see the original manuscript for a full list of contributors to the ManyEcoEvo project. -->
